---
title: "lab-07-simpsons.Rmd"
author: "GHZLAN ABDULLAH "
date: "19 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
view(Whickham)
```
    
Your answer: 
Observational , because we just watching the people and write the notes also we didn't involving in any situation or controlling any variables.

2.  
```{r observations}
nrow(Whickham)
```

Your answer;
1314, represent recorded participants' age, smoking status at baseline.

3.  

```{r number_variables}
ncol(Whickham)
```

Your answer:  
3, age (Numerical), Smoker and the outcome are categorical
age (integer) , smoker and outcome are (factor)


```{r var_type}
class(Whickham$age)
class(Whickham$smoker)
class(Whickham$outcome)
```

Your answer: 
age (Numerical) , smoker and outcome are categorical
age (integer)   , smoker and outcome are (factor)

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +
  geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +
  geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +
  geom_boxplot()
```


4. 
I expect the health will be worser and may be the person will be died after while, if he keeping smoke.

```{r plot_smokeVSincome_1}
 ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
 502+230
 230/732
```
smoker no  (732) :   31,4 (dead) >> (68,6) alive
smoker yes (582) :   23,8 (dead) >> (76,2) alive

i dose not expected this result because now the most died people not smoker .

6.

```{r age_cat}
Whickham <- Whickham%>% mutate (age_cat = case_when (age <= 44 ~ "18-44", age > 44. & age <= 64 ~ "45-64", age > 64 ~ "65+"))
```

7. 

```{r plot_smokeVSincome_3}
ggplot(data=Whickham, aes(x=outcome, y=smoker,color=smoker, fill=age_cat)) + geom_bar(stat="identity") +facet_wrap(~age_cat)

```

Knit, commit, and push to github.